\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[pdftex]{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{comment}
%%%% MIGHT TURN BACK ON WITH PXFONTS
%\usepackage{units,textcomp,xfrac}
%\usepackage{harvard}
\usepackage{amsmath, amsthm, amssymb}
\usepackage[colon]{natbib}
\usepackage{url}
\usepackage{fullpage}
\usepackage{bbm}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage{float}
\usepackage{tcolorbox}
\usepackage{verbatim}
\usepackage{pdfpages}
\usepackage{longtable}
\usepackage{lscape}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[english]{babel}
\usepackage{tabulary}
\usepackage{tabularx}
\usepackage{array}
\usepackage{sectsty}
%\usepackage{pdflscape}
\sectionfont{\large}
\usepackage{nicefrac}
\usepackage{rotating}
% \usepackage{floatrow} % added by MM for figure captions
% \usepackage{pxfonts}
% \usepackage{kpfonts}
\newcommand{\sups}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subs}[1]{\ensuremath{_{\textrm{#1}}}}

\usepackage{needspace}
%\usepackage[subtle]{savetrees}

%%% MM command for rounding macro numbers %%%%%%%%%%%%%%
\usepackage{numprint} 
\nprounddigits{2}
\npdecimalsign{.}
\npthousandsep{,}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\usepackage{subfig}
\usepackage{threeparttable}
%\usepackage{longtable}

% this allows you to use \sym from the stata-output tables
\newcommand{\sym}[1]{\rlap{#1}}

%%%%%%%%%%%%%%%%%%%%%%
%%%% APPENDIX STUFF
\usepackage[toc,page]{appendix}


%% COMMAND THAT ALLOWS YOU TO MAKE THE INDEPENDENCE SYMBOL
%% INDEPENDENCE SYMBOL ALSO REQUIRES GRAPHICX
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}


\floatstyle{plain}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}} % symbols-a4, p.106
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}} 

\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 

\let\footnotesize=\small
\let\titlesize=\small

\usepackage{geometry}
% \onehalfspacing
\geometry{verbose,letterpaper,tmargin=1.0in,bmargin=1in,lmargin=1in,rmargin=1in}
%\geometry{verbose,letterpaper,tmargin=1.5in,bmargin=1.5in,lmargin=1.5in,rmargin=1.5in}

\renewcommand*{\thetable}{\arabic{table}}

% \usepackage[colorlinks]{hyperref}
\PassOptionsToPackage{hyphens}{url}\usepackage[colorlinks]{hyperref}
\hypersetup{
  colorlinks=true,
  citecolor=Blue,
  linkcolor=Red,
  urlcolor=Magenta}

\begin{document}
\DeclareGraphicsExtensions{.pdf,.tif, .jpg}

\title{ \vspace{-1.5cm}
Beyond Crime Rates: How Did Public Safety in U.S. Cities Change in 2020?\thanks{Author order is randomized. Massenkoff: maxim.massenkoff@nps.edu, Chalfin: achalfin@sas.upenn.edu. We thank David Abrams, Jacob Kaplan, John MacDonald, Evan Rose, Yotam Shem-Tov and seminar participants at the Claremont Graduate University for helpful feedback on an earlier version of this paper. Any remaining errors are our own.
} }
%Beyond Crime Rates: Measuring Changes in the \emph{Risk} of Crime Victimization
%Beyond Crime Rates: How Did the Risk of Crime Victimization Actually Change During the COVID-19 Pandemic?
\author{
  {\large Maxim Massenkoff} \\ {\normalsize Naval Postgraduate School}  \\ 
  \and {\large Aaron Chalfin} \\ {\normalsize University of Pennsylvania and NBER}  
  }

\date{\vspace{.5cm}This version: \today}
% \date{\vspace{.5cm} November 5, 2020}

\maketitle
\thispagestyle{empty} 

% \vspace{-1cm}
% \begin{center}
% \textbf{PRELIMINARY AND INCOMPLETE; PLEASE DO NOT CITE OR CIRCULATE WITHOUT PERMISSION}
% \end{center}

\begin{abstract}
\setlength{\baselineskip}{13pt} %temporary set single space

\vspace{0.0in} %insert vertical space (in inches); can be in cm, mm
% Source for leading cause: https://www.cdc.gov/injury/images/lc-charts/leading_causes_of_death_by_age_group_2018_1100w850h.jpg

\noindent
This paper argues that changes in human activity during the COVID-19 pandemic led to an unusual divergence between crime rates and victimization risk in US cities. Most violent crimes declined during the pandemic. But analysis using data on activity shows that the \textit{risk} of street crime victimization was elevated throughout 2020. People in public spaces were 15-30 percent more likely to be robbed or assaulted. This increase is unlikely to be explained by changes in crime reporting or selection into outdoor activities by potential victims. Traditional crime rates may present a misleading view of recent changes in public safety.


\end{abstract}


\noindent


\thispagestyle{empty}\baselineskip1.5\baselineskip\newpage{}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\tableofcontents

\newpage
\setcounter{page}{1}

\begin{tcolorbox}
\textbf{Significance statement}  \\ 
The onset of the COVID-19 pandemic brought massive disruptions to economic and social life, including an unprecedented spike in homicides. Overall crime, however, was down. We resolve this apparent paradox by showing that after accounting for the fact that people were spending more time indoors in 2020, the risk of victimization in public actually increased. These recent changes in crime and activity provide a stark illustration of how conventional crime rates can fail to capture changes in public safety.

\end{tcolorbox}

\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The onset of the COVID-19 pandemic brought massive disruptions to economic and social life, including an unprecedented spike in homicides \citep{cdc2020hom}. Overall crime, however, was down, spurring a conventional narrative that gun violence has been an exception \citep{graham2021atlantic, nass2020trace} and that, contrary to public opinion, the pandemic has not had a deleterious impact on public safety \citep{ashby2020initial,abrams2021covid,koerth2020many}. This understanding, based on analyses of traditional crime data, informs an important discussion among researchers, public safety advocates, and journalists about recent changes in crime. But is it substantively correct?  

This paper illustrates how changes in human activity during the COVID-19 pandemic led to an unusual divergence between official crime rates and victimization risk. We focus, in particular, on public violence---assaults and robberies occurring in public or commercial spaces---which are often perpetrated by strangers and which drive an outsize share of fear among the public \citep{ferraro1995fear, pickett2012reconsidering, hanslmaier2013crime, ferraro2017measurement}. Beginning in March 2020, public violence declined by approximately 35\% as people responded to disease risk and mandated lockdowns by spending more time at home. These offenses remained 10-15\% below 2019 levels throughout the summer before returning to 2019 levels in the fall. But what happened to the risk of violent crime that people faced in while spending time in public spaces? Because our principal measure of public safety---the crime rate---is so limited, conventional analyses are unable to answer this question.\footnote{From its earliest inception, the per capita crime rate has always been intended to provide a rough proxy for victimization risk \citep{boggs1965urban, stipak1988alternatives, vaughan2020promise, ramos2021improving}. %Prior research has proposed various of ways of moving beyond crime rates to better capture victimization risk. 
We provide a brief history of prior thinking about how to compute an ``activity adjusted crime rate" in Appendix \ref{sec:appendixA}.}

%which is arguably one of the most fundamental questions about how urban life has changed during the pandemic.

\paragraph{Research strategy and key findings}

%We study the change in public victimization risk in U.S. cities between 2019 and 2020. Our core estimand is descriptive: how did risk change with the onset of the COVID pandemic? In theory, the ideal measure of risk would capture all public victimizations and the aggregate time that people spent in public in both years, yielding a measure of victimizations per hour in both periods. To estimate risk, we count crimes in public using police reports and victimization surveys. We measure public activity using newly available mobile location data and survey data on time use. 

Using national survey data, as well as information on the movement of people based on mobile location data in several large US cities, we study the risk of public victimization in 2020. Newly-available survey data from the American Time Use Survey as well as foot traffic data from Google, Apple, Facebook and Safegraph, show that time spent in public spaces fell markedly in March 2020 before slowly rebounding thereafter. While Americans eventually began to return to many of their normal activities, even at the end of 2020, time spent outdoors remained about 20\% lower than in 2019.

%Both the American Time Use Survey (ATUS) as well as foot traffic data from Google, Apple, Facebook and Safegraph, show that time spent in public spaces fell markedly in March 2020 before slowly rebounding thereafter. While Americans eventually began to return to many of their normal activities, even at the end of 2020, time spent outdoors remained about 20\% lower than in 2019. Crime data from large cities and a national survey also show distinct drops in victimization in the post-lockdown period.

We next compare the number of violent crimes to measures of time spent away from home in order to calculate the risk of street crime victimization. We estimate this as the count of violent crimes occurring in public spaces divided by a measure of time spent in public places. While analyses of traditional crime data show discrete drops in offending, we find that, in 2020, the risk of outdoor street crimes initially \emph{rose} by more than 40\% and was consistently between 10-25\% higher than it had been in 2019 through the remainder of the year.\footnote{For analyses of traditional crime data during the COVID-19 pandemic, see \citet{ashby2020initial, boman2020has, gerell2020minor, hodgkinson2020show, langton2021six, payne2020covid, piquero2020staying, shayegh2020staying, abrams2021covid, boman2021global, campedelli2021exploring, de2021druglords, nivette2021global}.}$^{,}$\footnote{These findings complement and greatly extend recent research which uses Google's COVID-19 Community Mobility Reports to assess post-pandemic changes in criminal opportunities in the United States \citep{lopez2021crime} and the United Kingdom \citep{halford2020crime}. While this research notes that criminal opportunities shifted in 2020, it does not utilize a pre-pandemic measure of mobility, does not distinguish between crimes that occurred in residential and public locations, does not address selection effects and ultimately, it does not construct a quantitative test of the change in victimization risk during the pandemic.}

The result is consistent across two independent analyses. The first analysis utilizes administrative crime data and location-based mobility data from Safegraph for New York City, Los Angeles and Chicago, the three largest cities in the United States. The second analysis relies entirely on national survey data from the National Crime Victimization Survey and the American Time Use Survey. Critically, the administrative data from NYC, Los Angeles and Chicago and the national survey data point to the same conclusion: that Americans were more likely to be robbed or assaulted while spending time in public after March 2020 than they had been previously. 

%Our principal estimand is the risk of victimization conditional on spending time in public. 
Our core estimand is descriptive, measuring the change in victimization risk that co-occurred with the onset of the COVID pandemic. While this paper is not designed to estimate a causal effect of the pandemic on public safety or account for all of the mechanisms through which the risk of crime increased, we consider two data artifacts that have implications for how these results are interpreted. First, it is possible that the reporting of crimes to law enforcement changed in 2020. Here, the available survey data shows no dramatic changes in the propensity of victims to report crimes to the police (see Appendix \ref{sec:reporting}). A second notable possibility is that the pool of potential victims may have changed. If in 2020, people selecting into outdoor activities---and who were thus at risk of a street crime---were those with a higher baseline risk of victimization, our aggregated estimates of crime risk might not speak to the change in risk for the typical person. We show in Appendix \ref{sec:selection} that compositional shifts among people out in public spaces imply no large changes in crime, and that 2020 victims were, if anything, at lower risk of victimization historically. These analyses suggest that our finding of increased risk does not appear to be an artifact of victim selection.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mobility}
Data on time spent in public spaces in 2019 and 2020 are presented in \autoref{fig:city_mobility_combined}. The figure presents trends using mobility data from Safegraph (panel A), Google (panel B), Apple (panel C), Facebook (panel D) as well as ATUS survey data (panel E). %\footnote{While all four mobility datasets are available in 2020, only the Safegraph data are available earlier, thus allowing us to observe changes in activity that account for seasonal variation. } 
For the mobility data, we restrict the sample to our three cities---New York, Los Angeles and Chicago---and aggregate across the cities to present information for the combined sample.\footnote{We present the same data, disaggregated by city in Appendix \ref{sec:appendix2} and note that patterns are extraordinarily similar across cities. While the time-path of the pandemic's impact differed widely among the three cities, the fact that mobility patterns are broadly similar suggests that these are driven to a greater extent by lockdowns or national trends than by localized health risks. }

All five sources of data indicate that outdoor foot traffic declined steeply at the beginning of the pandemic. In the Safegraph data, the total count of unique neighborhood visits across our three cities fell from an average of 220 million to 110 million from 2019 to 2020, a 50\% decrease.\footnote{The Google, Apple and Facebook data are more limited in that they are only available for 2020 and provided in terms of indexes relative to early 2020. Apple data is since January 13, 2020. The Google baseline is Jan 3–Feb 6, 2020 \citep{google_data}. Facebook baseline is from the end of February 2020 \citep{facebook_documentation}. Still, these data sources support a similar story.} The indexes in panels B, C and D suggest similar decreases. Apart from grocery shopping, the Google indexes in panel B decline by 60\% and remain 20-50\% lower throughout the year. The Apple travel indexes in panel C  fall by 60-80\% in April and show a sharper rebound---likely due to their indexing to a winter month. The Facebook index in panel D shows a sustained 20\% increase in staying put, a measure of the fraction of people staying in a small area for an entire day.

Finally, we consider national survey data from the ATUS in panel E, plotting the share of waking hours spent away from one's home by month in 2019 and 2020. In 2019, respondents reported spending approximately 7 of their waking hours away from home, a statistic which exhibits relatively little seasonal variation. By May 2020, this had fallen to 3 hours.\footnote{The ATUS was temporarily suspended between mid-March and mid-May 2020 due to pandemic-induced lockdowns.} Throughout the rest of 2020, time spent outside remained about 30\% lower than in the previous year. Though the precise magnitudes differ somewhat depending on the source of data used, the broad trends are similar.

\subsection{Victimization Risk}
We plot crimes known to law enforcement for our combined New York City-Los Angeles-Chicago sample in panel A of \autoref{fig:city_crime_combined}, presenting monthly counts of violent crimes separately for 2019 and 2020.\footnote{Results for individual cities can be found in Appendix \ref{sec:appendix2}. %An alternative set of results using a narrower definition of public locations is available in \autoref{fig:city_crime_alt_combined}.
} Public violence was approximately 10\% higher in the winter of 2020 than the winter of 2019 but began to fall steeply after March 2020. In April 2020, public violence was approximately one third lower than it had been in April 2019. Consistent with the seasonality documented in prior research \citep{andresen2013crime, jacob2007dynamics}, public violence rose during the spring and summer months in both years.  However, throughout the summer and into the fall months, these crimes remained between 5-20\% lower in 2020 than they had been in 2019. Aggregating across the entire pandemic period of 2020 (March through December), public violence was 19\% lower in 2020 than it had been in 2019.\footnote{A corresponding decline in violent victimization is also observed in survey data from the NCVS which recorded a 24\% decline between the 2019 and 2020 surveys.}

Panel B plots violent street crime risk per 100,000 people using Safegraph data to measure the change in activity. In March 2020, street crime victimization risk was roughly equal to risk in the previous year at approximately 5 violent street crimes per month per 100,000 person-visits to a community. %\footnote{This statistic cannot be directly compared to per capita crime measures generated from administrative data as the denominator uses a different proxy for exposure.} 
However, at the same time that street crimes fell by more than 30\% in April 2020, the \emph{risk} of street crime victimization rose by nearly 40\%. Victimization risk remained elevated throughout the summer---approximately 12\% higher in July 2020 than July 2019---and never fully returned to trend, even by the end of the year. For the March-December period, street crime victimization risk was approximately 14\% higher in 2020 than it had been in 2019, despite a 19\% decline in the incidence of public violence. 

Next, we present the same analysis using national survey data from the NCVS and the ATUS. Figure \ref{fig:ncvs_nat_ests}, Panel A plots the total number of outdoor violent crimes in the United States, using NCVS respondents living in metropolitan areas. The number of outoor violent crimes fell from approximately 1.6 million in 2019 to 1.3 million in 2020, a 17\% reduction. Panel B plots the number of violent victimizations occurring in public spaces divided by the number of hours spent in public spaces reported in the ATUS. While the number of crimes fell by 17\%, the \emph{risk} of public victimization rose by approximately 25\% in 2020---from 9 to  12 victimizations per million hours spent in public spaces. This analysis suggests that the increase in crime risk reported in \autoref{fig:city_crime_combined} is not unique to the three largest cities in the US, and provides further assurance that the city-specific findings are unlikely to be an artifact of the limitations of Safegraph mobility data. 

%Taken together, these results suggest that the increase in risk was not restricted to an aberration in gun violence. While gun violence rose by 47\% in the three largest cities in the United States and by around 30\% nationally, the risk of violent street crimes generally rose as well---by around 25\% nationally. %Though street crimes appear to be slightly less sensitive to the effects of the pandemic than shootings, the trends are much more similar after accounting for changes in criminal opportunities.

\subsection{Robustness and Extensions}
In Appendix \ref{sec:appendix1} we probe the robustness of our main results. We briefly summarize these analyses here. First, we reproduce \autoref{fig:city_crime_combined} using ATUS survey data to measure the change in mobility between 2019 and 2020 (Appendix Figure \ref{fig:atus_risk}). The estimate is even larger, suggesting a 32\% increase in risk over the sample period. Next, we consider an alternative definition of public violence, limiting the analysis to crimes that occur in outdoor locations as opposed to all non-residential locations and find similar results (Appendix Figure \ref{fig:city_crime_alt_combined}). We also consider whether the types of activities that people are engaged in while outside their homes may have shifted in 2020 towards activities that carry a higher level of inherent risk (Appendix Figure \ref{fig:atus_where}). While the amount of time spent in public decreased markedly in 2020, the proportion of time spent on each of several different activities did not show radical changes. The largest decrease is in the proportion of time in public spent walking, which would make activities in public lower risk.

Next, we quantify the uncertainty in our estimates of the change in risk for both the city-level and national analyses. We present bootstrapped confidence intervals for each of our estimates, along with computational details, in Appendix \ref{sec:inference}. For the city-level analysis using Safegraph data, we find that the increase in victimization risk is highly significant ($p$ $<$ 0.001), as the crime measures are not a sample and the city-wide decline in activity inferred from SafeGraph is precisely measured. The national survey data are sparser, and the 95\% confidence interval for our main estimate, the change in the risk of outside violent offenses in urban areas, is wide and includes zero ($p =0.136$). As an additional check, we show that the point estimate is similar and statistically significant at conventional levels as we broaden the type of crimes that are included.

Finally, %we consider whether our main result is likely to be a mechanical artifact of either the measurement of crime or changes in victim selection. Our principal conclusion is that the risk of public violence increased in the US in 2020. 
while an analysis of the causal effect of the pandemic is beyond the scope of this paper, we consider two potential explanations which, if true, might alter the policy implications of our findings. First, it is possible that the reporting of crimes to law enforcement changed in 2020. We compare the pooled 2015-2019 and 2020 waves of the NCVS and see little evidence that crime reporting changed appreciably in 2020, hovering around 40\% across periods.
We further note that our analysis of NCVS survey data includes reported as well as unreported crimes and is, as such, robust to this concern. This analysis is presented in Appendix \ref{sec:appendix_alternative}.
%Even taking literally the slight increase in reporting captured by the NCVS, the survey itself, which is not vulnerable to the same bias, captures a similar increase in risk as the city crime data. 

A second possibility is that different types of people spent time outdoors in 2020 than in 2019. If in 2020, people selecting into outdoor activities---and who were thus at risk of a street crime---were those with a higher baseline risk of victimization, our aggregated activity-adjusted estimates of crime risk would not necessarily imply an increase in risk for the typical person. We assess the importance of victim selection in three different ways (Appendix Section \ref{sec:selection}). First, in a reweighting exercise \citep{dinardo1996labor}, we use the ATUS and NCVS to calculate a counterfactual 2020 crime rate using observed activity changes and 2015-2019 victimization rates by age, race, gender and household income. We find that the counterfactual rates are nearly identical to actual crime, suggesting that selection into activity among higher-risk groups is not an important source of bias. Second, using microdata from New York City and Los Angeles, we directly examine changes in the demography of crime victims. If anything, crime victims in the police data had \textit{lower}, not higher, historical rates of victimization. Finally, to test whether the increase in risk is driven by people in higher crime neighborhoods spending relatively more time outside, we condition on Census block group fixed effects, finding that risk increased during the pandemic, even \emph{within a given community}. It remains possible that selection into outdoor activities may have occurred along other unobserved dimensions. But given the preeminence of age, gender, race and location in explaining variation in crime, these analyses narrow the scope for unobserved factors to have played a leading role in driving up victimization risk and support the idea that risk increased for the average person.

%Finally, to test whether the increase in risk is driven by people in high-risk neighborhoods spending relatively more time outside, we condition on Census block group fixed effects, finding that within-neighborhood risk also increased during the pandemic. These analyses support the idea that risk increased for the average person. %Each of these analyses are presented in Appendix \ref{sec:appendix_alternative}. %(Appendix \ref{sec:appendix_alternative}) 
%Thus the demographics and location of victims did not shift toward higher risk groups based on the recorded attributes in our data. However, we note that people may still have been selecting into activity within demographic groups and within neighborhoods, based on unobserved factors like risk preferences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We study changes in victimization risk using two different analyses. We begin with an analysis of administrative crime and mobility data from the three largest cities in the United States, computing an activity-adjusted measure of the risk of public violence. Next, we perform the same analysis using survey data from the National Crime Victimization Survey and the American Time Use Survey. In this section, we describe each of these data sources.

\subsection{Victimization Data}
There are two principal measures of crime in the United States. The first measure is derived from administrative data on crimes known to law enforcement. These data are compiled annually by the Federal Bureau of Investigation's Uniform Crime Reporting program and are increasingly made available directly to researchers by city planners in large US cities. The second measure of crime is derived from survey data from the National Crime Victimization Survey, an annual survey of a representative cross-section of American households. %We use each of these data sources, in turn.

\subsubsection{Administrative Crime data}
For our city-level analysis, crime is measured using publicly available microdata from New York City \citep{nyc_crime_data}, Los Angeles \citep{la_crime_data} and Chicago \citep{chicago_crime_data}. In each city, the microdata represent the universe of crimes known to law enforcement. For each crime, we observe the date, time, and type of crime. In New York City and Los Angeles, we also observe the age, gender and race of crime victims. %We use these data to address the possibility that the demographics of crime victims changed in 2020.

Critically, in all three cities we observe detailed information on the location of each crime, allowing us to identify whether the crime occurred inside a residence or in a public setting. In our primary analysis, we consider a crime to occur in a public space if it happened in any location that is not within the confines of a residential building or a connected property such as a yard or garage. Using this definition, public spaces include streets, parks and alleyways as well as commercial establishments and offices where people work. We also present alternative results where we count only streets, parks and alleyways as public spaces and confirm that the findings remain substantively unchanged (Appendix Section \ref{sec:alt_public}).   

\subsubsection{National Crime Victimization Survey}
Our national analysis measures crime victimization using data from the National Crime Victimization Survey (NCVS), a nationally representative repeat cross-sectional survey of American residents. The NCVS is widely regarded as the most accurate source of data on crime victimization and the reporting of crimes to law enforcement in the United States \citep{gutierrez2017silence}. The survey directs respondents to indicate if they have been the victim of a crime during the past six months. If so, respondents are asked to identify the types of crimes they have experienced and provide key details about each incident including the type of location in which the crime occurred. We use this information to focus on the subset of violent crimes that occurred in public spaces. Unlike the administrative crime sources, the survey data do not indicate the date or month of the victimization, so these analyses are limited to the annual level. For details about how we classify crimes as occurring in either public or residential locations in both the administrative data and the NCVS, see Appendix \ref{sec:safegraph_deets}, section 1.

Because rates of crime reporting to law enforcement may have changed in 2020, we present an additional analysis of reporting behavior using the 2015-2020 waves of the NCVS. Focusing on the subset of individuals living in metropolitan areas, we test whether victims became either more or less likely to report violent victimizations to law enforcement in 2020.

\subsection{Mobility data}
Our next task is to measure how the availability of potential street crime victims changed during the pandemic. In an ideal world we would observe the number of person-hours spent in public spaces and use this as a denominator against which to compare street crimes in 2019 and 2020. While activity cannot be captured perfectly using available data, we employ several data sources that allow us to construct proxy measures. In this section, we describe each of these data sources. 

% \footnote{The data have been used in several academic papers, listed \href{https://www.safegraph.com/publications/academic-research}{here}. Most applications are related to research on the COVID-19 pandemic, although see \citet{chen2019racial}.}

\subsubsection{Cell phone mobility data}
For our analysis of crime in New York City, Los Angeles and Chicago, our primary data source is SafeGraph's Neighborhood Patterns data, anonymized GPS data from cell phones with location services enabled \citep{sg_neighb}. These data provide a count of the number of unique people present in each Census block group in each month, disaggregated according to some broad features such as the Census block group that each visitor came from and whether a given day was a weekday versus a weekend day. %\footnote{A Census block group is a small enumeration area meant to contain 600-3,000 residents.} 
There are two distinct advantages of the Safegraph data.  First, the data provide a direct measurement of the movement of people throughout the day that does not rely on self-reports from survey data.  Second, microdata are available by neighborhood, which allows for city-level and even sub-city level analyses. 

% On this point, we note that SafeGraph has explored the potential selection bias of tracked users by comparing their geography, education, and household income to Census data, finding a high correlation, implying that the sample of users is representative of the population at the census block group level \citep{safegraph_bias}.
While the data allow us to identify foot traffic with considerable granularity, they are subject to limitations. In particular, the data come from a non-random sample of users using certain apps. If this sample exhibits differential trends in activity, we could mismeasure changes in risk. We test the reliability of the data by measuring its trends against four other known sources of mobility information. In addition, to account for the fact that the data is a sample all visitor counts are scaled by the inverse sampling probability in the home Census block group. More details about the data construction are in the Appendix \ref{sec:safegraph_deets}, section 2. 
 
As a check on the SafeGraph trends, we also use geolocation data gathered by \citet{google_data}, \citet{apple_data}, and \citet{facebook_data}.\footnote{These datasets have been used in several studies on COVID-19 and social distancing \citep[e.g.][]{cot2021mining, venter2020covid}; examples which emphasize methodology are \citet{ilin2021public} and \citet{arnal2020private}.} Each of the three sources uses a different technique to define mobility. The Facebook data consists of two daily indexes, provided at the county level, measuring how often people stayed in a given Bing tile (a 600 square meter area) and how many different tiles people visited \citep{facebook_mobility}. The Apple data measures direction requests for walking, driving, and transit compared to a baseline measured on January 13, 2020 \citep{cot2021mining}. The Google data show mobility trends across four different categories: grocery and pharmacy, parks, retail and recreation, residential, and workplaces \citep{google_mobility, cot2021mining}. The Apple data is based on all Apple Maps users, while Google and Facebook data is based on those with location services enabled. 

\subsubsection{The American Time Use Survey}
In our national analysis, we measure changes in time spent in public spaces using survey data from the American Time Use Survey (ATUS). The ATUS is designed to study how and where people living in the United States spend their waking hours. The survey has been administered annually to a nationally representative sample of between 20,000 and 40,000 American households since 2003 and asks respondents to carefully document how they spent their time during a recent day. Importantly, the data permit researchers to delineate between time spent at one's home versus elsewhere. Respondents are surveyed throughout the year except for a gap corresponding to April 2020. We use each individual's survey date to build a monthly repeated cross-section of time use, limiting the data to residents in metropolitan areas. Our measure of outdoor activity comes from aggregating time spent in all activities occurring outside of a home or yard.\footnote{While the ATUS does not ask respondents where they slept, it does record detailed data about where a respondent spent his or her waking hours.} 

\subsection{Methods}
The crime rate is the count of crimes divided by a measure of the population at risk. Our crime measure is the number of violent street crimes---assaults and robberies occurring in public---measured either using local administrative data or national survey data. The ideal denominator would be a direct measure of the amount of time individuals spent outdoors, so that a crime rate could be calculated in terms of person-hours spent in public. In theory, all victimizations and hours spent outside could be calculated for the full urban population in time $t$:
\begin{equation}
Risk_t = \frac{Crimes_t}{PersonHoursOutside_t}
\end{equation}

The change in $Risk$ across time periods is our main descriptive focus. We estimate the formulation in (1) two different ways. For our national analysis, we estimate the denominator in (1) using ATUS survey data, which allows us to directly compute a national estimate of the number of person-hours spent in public spaces in each year. The numerator is calculated using the NCVS. For our analysis of victimization risk in New York City, Los Angeles and Chicago, we measure the denominator using data from Safegraph. We focus on this mobility data source because it is available in both 2019 and 2020 and its measures have a direct interpretation.


The proxy that Safegraph provides for the denominator in (1) is a measure of the total number of visitors across neighborhoods, counting both residents and non-residents. This will increase by 1 for a given Census block group whenever a new person enters the area and remains there for at least a few minutes. The SafeGraph data are thus accurate proxies for our ideal denominator if overall time spent outdoors scales linearly with the number of unique census block groups visited. For instance, if Census block groups in New York City went from 15,000 unique daily visitors in 2019 to 10,000 unique daily visitors in 2020, this would register as a 33\% decline in outdoor activity in our data. However, this would overstate the decrease if people were spending the same amount of time outside but in a fewer number of neighborhoods, and it would be an underestimate if people were spending even less time outside when they were in their home neighborhoods. The alignment we see between SafeGraph and the other activity measures provides some assurance that idiosyncrasies in SafeGraph's data do not drive our results.




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Beginning in March 2020, the number of street crimes in America's cities declined sharply as people adapted to disease risk and mandated lockdowns by spending more time at home. In contrast to analyses of traditional crime data \citep{abrams2021covid}, we find that, in 2020, the risk of outdoor street crimes initially \emph{rose} by more than 40\% and was between 15\% and 30\% \emph{higher} than it had been in 2019 through the remainder of the year. These differences are unlikely to be a mechanical artifact of changes in crime reporting or differential selection into outdoor activity during the COVID-19 pandemic. 
The results suggest that the decline in overall crime in 2020 is an artifact of changes in criminal opportunities and that worrying changes in public safety were not restricted to a spike in homicides. %In contrast to the narrative in the popular media \citep{koerth2020many, nass2020trace}, the increase in homicides was not an aberration. 
These shifts mirror other troubling increases in motor vehicle fatalities \citep{stewart2022overview} and drug overdoses \citep{stephenson2020drug}.

%  Analyses of standard administrative crime data could present a misleading view of what has happened to public safety since the beginning of the COVID-19 pandemic.

Per capita crime rates are used to capture a notion of risk. But, as we argue, this shorthand became less reliable with the onset of the COVID-19 pandemic, especially with respect to public violence, crimes which require victims to spend time outside their homes and which drive an outsize share of fear. The data and methods we employ suggest a path forward in going beyond crime rates to more accurately measure risk, a long-time goal of crime researchers that is more attainable now due to recent advances in mobility data. By comparing public victimization to mobility, we can better understand how public safety varies across cities, among neighborhoods within cities, by time of day, or within demographic groups. Not limited to understanding the effects of the COVID-19 pandemic, better measurement can lead to improvements in policy, for example through more cost-effective deployment of police officers, social services, and public infrastructure investments such as street lighting.  

%---for example, more cost-effective deployment of police officers and social services and the enhanced targeting of investments in public infrastructure such as street lighting, garbage removal and the remediation of blighted properties. 

\newpage
\singlespacing
\bibliographystyle{chicago}
\bibliography{bib}


\input{figs}
\newpage 
%\clearpage 
%\input{tables}

\counterwithin{figure}{section}
%\counterwithin{table}{section}

\newpage 
\clearpage
\vspace*{90mm} \\
\begin{center}
\Huge{\textbf{Online Appendix}}
\end{center}

\newpage
\input{app}
%\input{app_figs}

\appendix 




\end{document}
